---
title: "Harvard Forest flux estimates"
output: html_document
date: "2024-02-28"
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggpubr)
library(ggthemes)
library(cowplot)
library(patchwork)
# library(calecopal)
library(here)
library(ggpmisc)
library(rstatix)
library(wesanderson)
```

## COMPASS paper: visualization of Harvard Forest deployment data

Read in csv if not already loaded in global environment ("HarvardForest_fluxestimates_fall2023.csv")

```{r}

# read in data:
# HF_fluxestimates <- read.csv("HarvardForest_fluxestimates_fall2023.csv", header = T)

# pal <- cal_palette(name = "superbloom2", n = 24, type = "continuous")

pal <- wes_palette(name = "Zissou1", n=24, type = "continuous")

HF_fluxestimates %>% 
  mutate(hour = hour(hour_of_obs)) %>%
  filter(fluxL_umolm2sec > -0.01) %>%
  filter(id != "fluxes_bot114" & id != "fluxes_bot112") %>%
  ggplot(aes(x = hour, y = fluxL_umolm2sec, color = as.factor(hour)))+
  # ggplot(aes(x = hour_of_obs, y = fluxL_umolm2sec))+
  # geom_point(size = 1, alpha = 0.5)+
  labs(x="hour of the day", 
       y=expression(paste("flux estimate, ",mu, "mol m"^-2, "s"^-1)))+
  geom_boxplot()+
  ylim(0,11)+
  ggtitle("Fluxbot Autochambers")+
  # geom_point()+
  facet_wrap(~stand)+
  theme_bw()+
  scale_color_manual(values = pal)+
  theme(legend.position = "none")
  
```

```{r}
HF_fluxestimates %>% 
  # mutate(hour = hour(hour_of_obs)) %>%
  # ggplot(aes(x = as.factor(hour), y = fluxL_umolm2sec, fill = stand))+
  # ggplot(aes(x = as.factor(hour), y = fluxL_umolm2sec, fill = stand))+
  ggplot(aes(x = fluxQ_umolm2sec, y = fluxL_umolm2sec))+
  geom_point(size = 1, alpha = 0.5)+
  labs(x="quadratic estimate", y="linear estimate")+
  # geom_boxplot()+
  # facet_wrap(~hour)+
  # geom_smooth(method = "lm")+
  theme_bw()
```

Plot to examine diurnal variability: 24, 48, and 72 hours of sampling during example of most consistent sampling period(s):

```{r}
#colors:
pal <- cal_palette(name = "superbloom3", n = 14, type = "continuous")

# 24 hours, Oct. 4th
HF_fluxestimates %>% 
  filter(hour_of_obs > "2023-10-09 00:00:59" & hour_of_obs < "2023-10-10 23:59:59") %>%
  filter(id != "fluxes_bot114" & id != "fluxes_bot112") %>%
  filter(fluxL_umolm2sec > -0.01) %>% 
  ggplot(aes(x = hour_of_obs, y = fluxL_umolm2sec, color = id))+
  geom_point(size = 2, alpha = 0.5)+
  geom_line()+
  facet_grid(~stand)+
  theme_bw()+
  ggtitle("October 9-10th, 2023: 48hrs total")+
  labs(x="", y=expression(paste("flux estimate, ",mu, "mol m"^-2, "s"^-1)))+
  theme(legend.position = "none")+
  scale_x_datetime(date_labels = "%H:%M")+
  theme(panel.spacing = unit(2, "lines"))+
  scale_color_manual(values = pal)
```

Explore raw data from chosen fluxbots:

```{r}

rawdata_fluxbot_106 %>%
  filter(Time > "2023-10-26 12:54:00" &
           Time < "2023-10-26 13:00:00") %>%
  filter(co2 < 6000) %>%
  # filter(id == "fluxes_bot114" | id == "fluxes_bot112") %>% 
  ggplot(aes(x = Time, y = co2))+
  geom_point(size = 1, alpha = 0.5)+
  # geom_line()+
  labs(x="", y=expression(paste(CO[2], " concentration (ppm)")))+
  # ggtitle("Fluxbot 13 raw CO2 detection")+
  theme_classic()
  
```

Double check some of the sketchy intervals

```{r}

# Example usage of visualize_interval_regressions function
# Specify the start and end timestamps of the interval you want to visualize
interval_start <- as.POSIXct("2023-10-26 12:56:00", tz = "America/New_York")
interval_end <- as.POSIXct("2023-10-26 12:59:59", tz = "America/New_York")

test <- convert_co2_to_mass(rawdata_fluxbot_106, interval_start, interval_end)


rawdata_fluxbot_106 %>% 
  filter(Time > "2023-10-26 12:55:59" & Time < "2023-10-26 13:00:00") %>% 
  ggplot(aes(x=Time, y=test))+
  # ggplot(aes(x=Time, y=co2))+
  # ggplot(aes(x=Time, y=pressure))+
  geom_point()+
  # ylab("CO2 concentration (PPM)")+
  # ylab("CO2 mass (kg)")+
  theme_classic()
```

## Harvard Forest autochamber flux estimates: from Ashley and Corey, April 2024

upload HF autochamber data:

```{r}
# site 1 (aka "unhealthy" hemlock plot)
HFauto_unhealthy <- read.csv("site1_October2023_man.csv", header = T)
# site 2 (aka "healthy" hemlock plot)
HFauto_healthy <- read.csv("site2_October2023_man.csv", header = T)
# add stand info:
HFauto_unhealthy <- HFauto_unhealthy %>% 
  mutate(stand = "unhealthy")
# add stand info:
HFauto_healthy <- HFauto_healthy %>% 
  mutate(stand = "healthy")

# combine:
HFauto_fluxestimates <- rbind(HFauto_unhealthy, HFauto_healthy)

# visualize:
pal <- cal_palette(name = "superbloom2", n = 24, type = "continuous")

HFauto_fluxestimates %>% 
  # filter(rs > -0.01 & rs < 5) %>%
  filter(rs > -0.01 & rs < 15) %>% 
  ggplot(aes(x = hour, y = rs, color = as.factor(hour)))+
  # ggplot(aes(x = datetime, y = rs))+
  labs(x="hour of the day", 
       y=expression(paste("flux estimate, ",mu, "mol m"^-2, "s"^-1)))+
  geom_boxplot()+  ylim(0,11)+
  ggtitle("Harvard Forest Autochambers")+
  # geom_point()+
  facet_wrap(~stand)+
  theme_bw()+
  scale_color_manual(values = pal)+
  theme(legend.position = "none")
```

Combine the two datasets by stand and time:

```{r}
# summary stats by stand, autochamber array:
HFauto_fluxestimates %>%
  group_by(stand) %>% 
  summarise(mean_flux = mean(rs, na.rm = TRUE),
            min_flux = min(rs, na.rm = TRUE),
            max_flux = max(rs, na.rm = TRUE),
            sd_flux = sd(rs, na.rm = TRUE),
            tenth_quantile = quantile(rs, probs = .1, na.rm = TRUE),
            ninetieth_quantile = quantile(rs, probs = .9, na.rm = TRUE),
            variance = var(rs, na.rm = TRUE))
```

```{r}
# summary stands by stand, fluxbot array:
HF_fluxestimates %>%
  # filter(fluxL_umolm2sec > -1) %>% 
  group_by(stand) %>% 
  summarise(mean_flux = mean(fluxL_umolm2sec, na.rm = TRUE),
            min_flux = min(fluxL_umolm2sec, na.rm = TRUE),
            max_flux = max(fluxL_umolm2sec, na.rm = TRUE),
            sd_flux = sd(fluxL_umolm2sec, na.rm = TRUE),
            tenth_quantile = quantile(fluxL_umolm2sec, probs = .1, na.rm = TRUE),
            ninetieth_quantile = quantile(fluxL_umolm2sec, probs = .9, na.rm = TRUE),
            variance = var(fluxL_umolm2sec))
```

XY plot of hourly means from each array, and associated R2:

```{r}
# autochamber:
HFauto_hrlymeans <- HFauto_fluxestimates %>% 
  group_by(hour, stand) %>% 
  summarize(mean_hrly_auto = mean(rs, na.rm = TRUE))
# bots:
HFbots_hrlymeans <- HF_fluxestimates %>% 
  mutate(hour = lubridate::hour(hour_of_obs)) %>% 
  group_by(hour, stand) %>% 
  summarize(mean_hrly_bots = mean(fluxL_umolm2sec, na.rm = TRUE))

# bind:
hrlymeans_bystand <- merge(HFauto_hrlymeans, HFbots_hrlymeans, by = c("hour", "stand"))

# plot:
hrlymeans_bystand %>% 
  ggplot(aes(x=mean_hrly_auto, y=mean_hrly_bots, color = "pink"))+
  geom_point()+
  geom_smooth(method = "lm", se = TRUE, color = "black")+
  stat_poly_eq(aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
               formula = y~x,
               parse = TRUE,
               label.x = "right",
               label.y = "top")+
  labs(x = "mean hourly flux, HF autochambers", y = "mean hourly flux, Fluxbots")+
  facet_wrap(~stand)+
  theme_bw()+
  theme(
    legend.position = "none",
        axis.title.x = element_text(face = "bold", color = "darkred"),
        axis.title.y = element_text(face = "bold", color = "blue"))
```

Just all, not by stand:

```{r}
# autochamber:
HFauto_hrlymeansall <- HFauto_fluxestimates %>% 
  group_by(hour) %>% 
  summarize(mean_hrly_auto = mean(rs, na.rm = TRUE))
# bots:
HFbots_hrlymeansall <- HF_fluxestimates %>% 
  filter(fluxL_umolm2sec > -1) %>% 
  mutate(hour = lubridate::hour(hour_of_obs)) %>% 
  group_by(hour) %>% 
  summarize(mean_hrly_bots = mean(fluxL_umolm2sec, na.rm = TRUE))

# bind:
hrlymeans_all <- merge(HFauto_hrlymeansall, HFbots_hrlymeansall, by = c("hour"))

# plot:
hrlymeans_all %>% 
  ggplot(aes(x=mean_hrly_auto, y=mean_hrly_bots, color = "pink"))+
  geom_point()+
  geom_smooth(method = "lm", se = TRUE, color = "black")+
  stat_poly_eq(aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
               formula = y~x,
               parse = TRUE,
               label.x = "right",
               label.y = "top")+
  labs(x = "mean hourly flux, HF autochambers", y = "mean hourly flux, Fluxbots")+
  theme_bw()+
  theme(
    legend.position = "none",
        axis.title.x = element_text(face = "bold", color = "darkred"),
        axis.title.y = element_text(face = "bold", color = "blue"))
```

Same, but for every hour of the deployment (aka averaged across bots, not across the deployment period):

```{r}
# autochambers:
HFauto_fluxestimates <- HFauto_fluxestimates %>% 
  mutate(date_time = gsub("T", " ", datetime)) %>% 
  mutate(date_time = as.POSIXct(date_time, format = "%Y-%m-%d %H:%M:%S")) %>% 
  mutate(date = as_date(date_time))

# bin observations into buckets: average by time of day, "dawn", "morning", "midday", "afternoon", "dusk", "night"
HFauto_timeofday <- HFauto_fluxestimates %>% 
  mutate(timeofday = case_when(hour > 4 & hour < 8 ~ "dawn",
                                hour > 7 & hour < 11 ~ "morning",
                                hour > 10 & hour < 15 ~ "midday", 
                                hour > 14 & hour < 17 ~ "afternoon",
                                hour > 16 & hour < 20 ~ "dusk",
                                TRUE ~ "night")) %>% 
  group_by(date, timeofday) %>% 
  summarize(mean_timeofday_auto = mean(rs, na.rm = TRUE),
            stand = stand) %>% 
  unite("date_timeofday", date:timeofday, sep = ", ", remove = FALSE)

# fluxbots: remove values less than -1 --
HFbots_timeofday <- HF_fluxestimates %>% 
  mutate(date = as_date(hour_of_obs)) %>% 
  filter(fluxL_umolm2sec > -1) %>% 
  # pull "hour" out:
  mutate(hour = hour(hour_of_obs)) %>% 
  mutate(timeofday = case_when(hour > 4 & hour < 8 ~ "dawn",
                                hour > 7 & hour < 11 ~ "morning",
                                hour > 10 & hour < 15 ~ "midday", 
                                hour > 14 & hour < 17 ~ "afternoon",
                                hour > 16 & hour < 20 ~ "dusk",
                                TRUE ~ "night")) %>% 
  group_by(date, timeofday) %>% 
  summarize(mean_timeofday_bots = mean(fluxL_umolm2sec, na.rm = TRUE),
            stand = stand) %>% 
  unite("date_timeofday", date:timeofday, sep = ", ", remove = FALSE)

# combine:
means_timeofday <- merge(HFauto_timeofday, HFbots_timeofday, by = "date_timeofday")
```

Plot: by healthy, unhealthy plots

```{r}
legend_title <- "time of day"
means_timeofday$timeofday.x <- factor(means_timeofday$timeofday.x,
                                      levels = c("dawn", "morning", "midday", "afternoon", "dusk", "night"))

means_timeofday %>% 
  ggplot(aes(x=mean_timeofday_auto, y=mean_timeofday_bots, color = timeofday.x, fill = timeofday.x))+
  geom_point(color = "black", pch = 21)+
  geom_smooth(aes(color=timeofday.x), method = "lm", se = FALSE, linewidth = 0.5)+
  scale_fill_manual(legend_title, values = c("white", "lemonchiffon2", "yellow2", "goldenrod3", "steelblue4", "grey26"))+
  scale_color_manual(legend_title, values = c("white", "lemonchiffon2", "yellow2", "goldenrod3", "steelblue4", "grey26"))+
  stat_poly_eq(aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
               formula = y~x,
               parse = TRUE,
               label.x = "right",
               label.y = "bottom",
               )+
  labs(x = "mean flux, HF autochambers", y = "mean flux, Fluxbots")+
  # facet_wrap(~stand.x)+
  theme_dark()

```

Plot: by time of day

```{r}
legend_title <- "plot condition"
means_timeofday$stand.x <- factor(means_timeofday$stand.x,
                                      levels = c("healthy", "unhealthy"))

means_timeofday %>% 
  ggplot(aes(x=mean_timeofday_auto, y=mean_timeofday_bots, color = stand.x, fill = stand.x))+
  geom_point(color = "black", pch = 21)+
  geom_smooth(aes(color=stand.x), method = "lm", se = FALSE, linewidth = 0.5)+
  scale_fill_manual(legend_title, values = c("darkolivegreen3", "lightsalmon1"))+
  scale_color_manual(legend_title, values = c("darkolivegreen3", "lightsalmon1"))+
  stat_poly_eq(aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
               formula = y~x,
               parse = TRUE,
               label.x = "right",
               label.y = "bottom",
               vstep = .1,
               size = 3
               )+
  labs(x = "mean flux, HF autochambers", y = "mean flux, Fluxbots")+
  facet_wrap(~timeofday.x)+
  theme_dark()

```

Plot means of data across time:

```{r}
a <- means_timeofday %>% 
  ggplot(aes(x=timeofday.x, y=mean_timeofday_auto, fill=stand.x))+
  geom_boxplot()+
  theme_classic2()+
  theme(legend.position = c(.2, .1),
        legend.text = element_text(size = 8), legend.title = element_blank())+
  ylim(0,6)+
  labs(x="", y="average flux by time of day", title = "HF autochambers")+
  scale_fill_manual(legend_title, values = c("darkolivegreen3", "lightsalmon1"))

b <- means_timeofday %>% 
  ggplot(aes(x=timeofday.x, y=mean_timeofday_bots, fill=stand.x))+
  geom_boxplot()+
  theme_classic2()+
  theme(legend.position = c(.2, .1),
        legend.text = element_text(size = 8), legend.title = element_blank())+
  ylim(0,6)+
  labs(x="", y="average flux by time of day", title = "Fluxbots")+
  scale_fill_manual(legend_title, values = c("darkolivegreen4", "lightsalmon3"))

ggarrange(a,b)

```

redo the summary stats with time of day associated:

```{r}
HF_fluxestimates <- HF_fluxestimates %>% 
  mutate(date = as_date(hour_of_obs)) %>% 
  # pull "hour" out:
  mutate(hour = hour(hour_of_obs)) %>% 
  # assign time of day to each observation:
  mutate(timeofday = case_when(hour > 4 & hour < 8 ~ "dawn",
                                hour > 7 & hour < 11 ~ "morning",
                                hour > 10 & hour < 15 ~ "midday", 
                                hour > 14 & hour < 17 ~ "afternoon",
                                hour > 16 & hour < 20 ~ "dusk",
                                TRUE ~ "night"))

HFauto_fluxestimates <- HFauto_fluxestimates %>% 
  # mutate(date = as_date(hour_of_obs)) %>% 
  # pull "hour" out:
  # mutate(hour = hour(hour_of_obs)) %>% 
  # assign time of day to each observation:
  mutate(timeofday = case_when(hour > 4 & hour < 8 ~ "dawn",
                                hour > 7 & hour < 11 ~ "morning",
                                hour > 10 & hour < 15 ~ "midday", 
                                hour > 14 & hour < 17 ~ "afternoon",
                                hour > 16 & hour < 20 ~ "dusk",
                                TRUE ~ "night"))

# summary stats by time of day:
HF_fluxestimates$timeofday <- factor(HF_fluxestimates$timeofday,
                                      levels = c("dawn", "morning", "midday", "afternoon", "dusk", "night"))
HF_fluxestimates %>%
  filter(fluxL_umolm2sec > -1) %>%
  group_by(timeofday) %>% 
  summarise(mean_flux = mean(fluxL_umolm2sec, na.rm = TRUE),
            min_flux = min(fluxL_umolm2sec, na.rm = TRUE),
            max_flux = max(fluxL_umolm2sec, na.rm = TRUE),
            sd_flux = sd(fluxL_umolm2sec, na.rm = TRUE),
            tenth_quantile = quantile(fluxL_umolm2sec, probs = .1, na.rm = TRUE),
            ninetieth_quantile = quantile(fluxL_umolm2sec, probs = .9, na.rm = TRUE),
            variance = var(fluxL_umolm2sec))

HFauto_fluxestimates$timeofday <- factor(HFauto_fluxestimates$timeofday,
                                      levels = c("dawn", "morning", "midday", "afternoon", "dusk", "night"))
HFauto_fluxestimates %>%
  filter(rs < 15) %>%
  group_by(timeofday) %>% 
  summarise(mean_flux = mean(rs, na.rm = TRUE),
            min_flux = min(rs, na.rm = TRUE),
            max_flux = max(rs, na.rm = TRUE),
            sd_flux = sd(rs, na.rm = TRUE),
            tenth_quantile = quantile(rs, probs = .1, na.rm = TRUE),
            ninetieth_quantile = quantile(rs, probs = .9, na.rm = TRUE),
            variance = var(rs, na.rm = TRUE))
```

Plot non-meaned data across time:

```{r}
c <- HFauto_fluxestimates %>% 
  filter(rs < 15) %>% 
  ggplot(aes(x=timeofday, y=rs, fill=stand))+
  geom_boxplot()+
  theme_classic2()+
  theme(legend.position = c(.2, .9),
        legend.text = element_text(size = 8), legend.title = element_blank())+
  ylim(0,12)+
  labs(x="", y="spread of flux estimates by time of day", title = "HF autochambers")+
  scale_fill_manual(legend_title, values = c("darkolivegreen3", "lightsalmon1"))

d <- HF_fluxestimates %>% 
  filter(fluxL_umolm2sec > 0) %>% 
  ggplot(aes(x=timeofday, y=fluxL_umolm2sec, fill=stand))+
  geom_boxplot()+
  theme_classic2()+
  theme(legend.position = c(.2, .9),
        legend.text = element_text(size = 8), legend.title = element_blank())+
  ylim(0,12)+
  labs(x="", y="spread of flux estimates by time of day", title = "Fluxbots")+
  scale_fill_manual(legend_title, values = c("darkolivegreen4", "lightsalmon3"))

ggarrange(c,d)

```

Simple statistics comparing healthy, unhealthy plots:

```{r}
# using library rstatix for tidy-compatible testing:

# HF AUTOCHAMBERS
t <- HFauto_fluxestimates %>% 
  filter(rs < 15) %>% 
  t_test(rs ~ stand, ref.group = "healthy", detailed = TRUE)
t
# healthy sample size: 6121, healthy mean: 3.048552
# unhealthy sample size: 7750, unhealthy mean: 2.33636
# statistic = 34.98524	
# df = 8264.074		
# p = 3.45e-250

# FLUXBOTS
t2 <- HF_fluxestimates %>% 
  filter(fluxL_umolm2sec > 0) %>% 
  t_test(fluxL_umolm2sec ~ stand, ref.group = "healthy", detailed = TRUE)
t2
# healthy sample size: 3639, healthy mean: 2.33272
# unhealthy sample size: 3366, unhealthy mean: 2.541244
# statistic = -6.67173		
# df = 6674.28	
# p = 2.73e-11	
```
