---
title: "Harvard Forest flux estimates"
output: html_document
date: "2025-06-02"
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggpubr)
library(ggthemes)
library(cowplot)
library(patchwork)
#library(calecopal)
library(wesanderson)
library(here)
library(ggpmisc)
library(rstatix)
```

## Comparison of HF autochambers, fluxbots array:

Read in data from fluxbots, and autochamber array data calculated using fluxbot R function:

```{r}

# met-data:
met_data <- read.csv("https://harvardforest.fas.harvard.edu/data/p00/hf001/hf001-10-15min-m.csv")

# fluxbot estimates:
HF_fluxestimates <- read.csv("/Users/elizabethforbes/Documents/HarvardForest_fluxbot/manuscript_code/HarvardForest_fluxestimates_fall2023.csv", header = TRUE)

pal <- wes_palette(name = "Zissou1", n = 24, type = "continuous")

# hour(ymd_hms(HF_fluxestimates$hour_of_obs))

HF_fluxestimates %>% 
  mutate(hour = hour(ymd_hms(hour_of_obs))) %>%
  filter(fluxL_umolm2sec > -0.01) %>%
  filter(id != "fluxes_bot114" & id != "fluxes_bot112") %>%
  ggplot(aes(x = hour, y = fluxL_umolm2sec, color = as.factor(hour)))+
  # ggplot(aes(x = hour_of_obs, y = fluxL_umolm2sec))+
  # geom_point(size = 1, alpha = 0.5)+
  labs(x="hour of the day", 
       y=expression(paste("flux estimate, ",mu, "mol m"^-2, "s"^-1)))+
  geom_boxplot()+
  ylim(0,11)+
  ggtitle("Fluxbot Autochambers")+
  # geom_point()+
  facet_wrap(~stand)+
  theme_bw()+
  scale_color_manual(values = pal)+
  theme(legend.position = "none")
  
```

Use custom function (filter_iqr) to filter out those flux estimates that fall outside the upper and lower quartile of the data's distribution:

```{r}
HF_fluxestimates <- HF_fluxestimates %>%
  filter(final_flux_umolm2sec >= 0) 

source("filter_iqr.R")
HF_fluxestimates<- HF_fluxestimates %>% filter_iqr('final_flux_umolm2sec') 

# realizing there are some instances where "hour of obs" was not calculated to include a time if it's rounding up from midnight...
HF_fluxestimates <- HF_fluxestimates %>% 
    mutate(hour_of_obs = {
    dt <- as.POSIXct(start_timestamp, format="%Y-%m-%d %H:%M:%S", tz="America/New_York")
    minutes <- minute(dt)
    if_else(minutes >= 30, 
            ceiling_date(dt, unit="hour"),
            floor_date(dt, unit="hour"))
  })

HF_fluxestimates %>% 
  # mutate(hour = hour(hour_of_obs)) %>%
  # ggplot(aes(x = as.factor(hour), y = fluxL_umolm2sec, fill = stand))+
  # ggplot(aes(x = as.factor(hour), y = fluxL_umolm2sec, fill = stand))+
  ggplot(aes(y = fluxQ_umolm2sec, x = fluxL_umolm2sec))+
  geom_point(size = 1, alpha = 0.5)+
  labs(x="quadratic estimate", y="linear estimate")+
  # geom_boxplot()+
  # facet_wrap(~hour)+
   geom_smooth(method = "lm")+
    geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed", size = 1) +  
  theme_bw()
```

Plot to examine diurnal variability: 24, 48, and 72 hours of sampling during example of most consistent sampling period(s):

```{r}
#colors:
#pal <- cal_palette(name = "superbloom3", n = 14, type = "continuous")

# 24 hours, Oct. 4th
HF_fluxestimates %>% 
  filter(hour_of_obs > "2023-10-09 00:00:59" & hour_of_obs < "2023-10-10 23:59:59") %>%
  filter(id != "fluxes_bot114" & id != "fluxes_bot112") %>%
  filter(fluxL_umolm2sec > -0.01) %>% 
  ggplot(aes(x = hour_of_obs, y = final_flux_umolm2sec, color = id))+
  geom_point(size = 2, alpha = 0.5)+
  geom_line()+
  facet_grid(~stand)+
  theme_bw()+
  ggtitle("October 9-10th, 2023: 48hrs total")+
  labs(x="", y=expression(paste("flux estimate, ",mu, "mol m"^-2, "s"^-1)))+
  theme(legend.position = "none")+
  #scale_x_datetime(date_labels = "%H:%M")+
  theme(panel.spacing = unit(2, "lines"))+
  scale_color_manual(values = pal)
```

Explore raw data from chosen fluxbots:

```{r}
# 
# rawdata_fluxbot_106 %>%
#   filter(Time > "2023-10-26 12:54:00" &
#            Time < "2023-10-26 13:00:00") %>%
#   filter(co2 < 6000) %>%
#   # filter(id == "fluxes_bot114" | id == "fluxes_bot112") %>% 
#   ggplot(aes(x = Time, y = co2))+
#   geom_point(size = 1, alpha = 0.5)+
#   # geom_line()+
#   labs(x="", y=expression(paste(CO[2], " concentration (ppm)")))+
#   # ggtitle("Fluxbot 13 raw CO2 detection")+
#   theme_classic()
#   
```

Double check some of the sketchy intervals

```{r}

# # Example usage of visualize_interval_regressions function
# # Specify the start and end timestamps of the interval you want to visualize
# interval_start <- as.POSIXct("2023-10-26 12:56:00", tz = "America/New_York")
# interval_end <- as.POSIXct("2023-10-26 12:59:59", tz = "America/New_York")
# 
# test <- convert_co2_to_mass(rawdata_fluxbot_106, interval_start, interval_end)
# 
# 
# rawdata_fluxbot_106 %>% 
#   filter(Time > "2023-10-26 12:55:59" & Time < "2023-10-26 13:00:00") %>% 
#   ggplot(aes(x=Time, y=test))+
#   # ggplot(aes(x=Time, y=co2))+
#   # ggplot(aes(x=Time, y=pressure))+
#   geom_point()+
#   # ylab("CO2 concentration (PPM)")+
#   # ylab("CO2 mass (kg)")+
#   theme_classic()
```

## Harvard Forest autochamber flux estimates

Calculated using the fluxbot code, and weather station pressure data.

```{r}
HFauto_fluxestimates <- read.csv("HFarray_fluxes_calculatedusingfluxbotcode.csv", 
                                   header = T)

# add stand column/info: site 1 = unhealthy (chambers 1-6), site 2 = healthy (chambers 7-12)
HFauto_fluxestimates <- HFauto_fluxestimates %>% 
  mutate(stand = case_when(
    autochamber == 1 | autochamber == 2 | autochamber == 3 | 
      autochamber == 4 | autochamber == 5 | autochamber == 6 ~ "unhealthy",
    autochamber == 7 | autochamber == 8 | autochamber == 9 | 
      autochamber == 10 | autochamber == 11 | autochamber == 12 ~ "healthy"
  )) %>% 
  rename(id = autochamber)

# add hour info: this rounds DOWN to the nearest hour (current hour), which is something to consider; could use a custom function.
HFauto_fluxestimates <- HFauto_fluxestimates %>% 
  # custom approach to round to the nearest hour while maintaining dates and times:
  mutate(hour_of_obs = {
    dt <- as.POSIXct(start_timestamp, format="%Y-%m-%d %H:%M:%S", tz="America/New_York")
    minutes <- minute(dt)
    if_else(minutes >= 30, 
            ceiling_date(dt, unit="hour"),
            floor_date(dt, unit="hour"))
  })

```

Combine the two datasets by stand and time:

```{r}
# summary stats by stand, autochamber array:
HFauto_fluxestimates %>%
  group_by(stand) %>% 
  summarise(mean_flux = mean(fluxL_umolm2sec, na.rm = TRUE),
            min_flux = min(fluxL_umolm2sec, na.rm = TRUE),
            max_flux = max(fluxL_umolm2sec, na.rm = TRUE),
            sd_flux = sd(fluxL_umolm2sec, na.rm = TRUE),
            tenth_quantile = quantile(fluxL_umolm2sec, probs = .1, na.rm = TRUE),
            ninetieth_quantile = quantile(fluxL_umolm2sec, probs = .9, na.rm = TRUE),
            variance = var(fluxL_umolm2sec, na.rm = TRUE))
```

```{r}
# summary stands by stand, fluxbot array:
HF_fluxestimates %>%
  # filter(fluxL_umolm2sec > -1) %>% 
  group_by(stand) %>% 
  summarise(mean_flux = mean(fluxL_umolm2sec, na.rm = TRUE),
            min_flux = min(fluxL_umolm2sec, na.rm = TRUE),
            max_flux = max(fluxL_umolm2sec, na.rm = TRUE),
            sd_flux = sd(fluxL_umolm2sec, na.rm = TRUE),
            tenth_quantile = quantile(fluxL_umolm2sec, probs = .1, na.rm = TRUE),
            ninetieth_quantile = quantile(fluxL_umolm2sec, probs = .9, na.rm = TRUE),
            variance = var(fluxL_umolm2sec))
```

XY plot of hourly means from each array, and associated R2:

```{r}
# autochamber:
HFauto_hrlymeans <- HFauto_fluxestimates %>% 
  mutate(hour = hour(ymd_hms(hour_of_obs))) %>% 
  group_by(hour, stand) %>% 
  summarize(mean_hrly_auto = mean(fluxL_umolm2sec, na.rm = TRUE))
  
# bots:
HFbots_hrlymeans <- HF_fluxestimates %>% 
  mutate(hour = hour(ymd_hms(hour_of_obs))) %>% 
  group_by(hour, stand) %>% 
  summarize(mean_hrly_bots = mean(fluxL_umolm2sec, na.rm = TRUE))

# bind:
hrlymeans_bystand <- merge(HFauto_hrlymeans, HFbots_hrlymeans, by = c("hour", "stand"))

# plot:
hrlymeans_bystand %>% 
  ggplot(aes(x=mean_hrly_auto, y=mean_hrly_bots, color = "pink"))+
  geom_point()+
  geom_smooth(method = "lm", se = TRUE, color = "black")+
  stat_poly_eq(aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
               formula = y~x,
               parse = TRUE,
               label.x = "right",
               label.y = "top")+
  labs(x = "mean hourly flux, HF autochambers", y = "mean hourly flux, Fluxbots")+
  facet_wrap(~stand)+
  theme_bw()+
  theme(
    legend.position = "none",
        axis.title.x = element_text(face = "bold", color = "darkred"),
        axis.title.y = element_text(face = "bold", color = "blue"))+
    geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed", size = 1) +
  xlim(2,3)+ylim(2,3)
```

Just all, not by stand:

```{r}
# autochamber:
HFauto_hrlymeansall <- HFauto_fluxestimates %>% 
  mutate(hour = hour(ymd_hms(hour_of_obs))) %>% 
  group_by(hour) %>% 
  summarize(mean_hrly_auto = mean(fluxL_umolm2sec, na.rm = TRUE))
# bots:
HFbots_hrlymeansall <- HF_fluxestimates %>% 
  filter(fluxL_umolm2sec > -1) %>% 
  mutate(hour = hour(ymd_hms(hour_of_obs))) %>% 
  group_by(hour) %>% 
  summarize(mean_hrly_bots = mean(fluxL_umolm2sec, na.rm = TRUE))

# bind:
hrlymeans_all <- merge(HFauto_hrlymeansall, HFbots_hrlymeansall, by = c("hour"))

# plot:
hrlymeans_all %>% 
  ggplot(aes(x=mean_hrly_auto, y=mean_hrly_bots, color = "pink"))+
  geom_point()+
  geom_smooth(method = "lm", se = TRUE, color = "black")+
  stat_poly_eq(aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
               formula = y~x,
               parse = TRUE,
               label.x = "right",
               label.y = "top")+
  labs(x = "mean hourly flux, HF autochambers", y = "mean hourly flux, Fluxbots")+
  theme_bw()+
  theme(
    legend.position = "none",
        axis.title.x = element_text(face = "bold", color = "darkred"),
        axis.title.y = element_text(face = "bold", color = "blue"))
```

Same, but for every hour of the deployment (aka averaged across bots, not across the deployment period):

```{r}
# autochambers:
HFauto_fluxestimates <- HFauto_fluxestimates %>% 
  mutate(date_time = gsub("T", " ", datetime)) %>% 
  mutate(date_time = as.POSIXct(date_time, format = "%Y-%m-%d %H:%M:%S")) %>% 
  mutate(date = as_date(date_time))

# bin observations into buckets: average by time of day, "dawn", "morning", "midday", "afternoon", "dusk", "night"
HFauto_timeofday <- HFauto_fluxestimates %>% 
  mutate(timeofday = case_when(hour > 4 & hour < 8 ~ "dawn",
                                hour > 7 & hour < 11 ~ "morning",
                                hour > 10 & hour < 15 ~ "midday", 
                                hour > 14 & hour < 17 ~ "afternoon",
                                hour > 16 & hour < 20 ~ "dusk",
                                TRUE ~ "night")) %>% 
  group_by(date, timeofday) %>% 
  summarize(mean_timeofday_auto = mean(rs, na.rm = TRUE),
            stand = stand) %>% 
  unite("date_timeofday", date:timeofday, sep = ", ", remove = FALSE)

# fluxbots: remove values less than -1 --
HFbots_timeofday <- HF_fluxestimates %>% 
  mutate(date = as_date(hour_of_obs)) %>% 
  filter(fluxL_umolm2sec > -1) %>% 
  # pull "hour" out:
  mutate(hour = hour(ymd_hms(hour_of_obs))) %>% 
  mutate(timeofday = case_when(hour > 4 & hour < 8 ~ "dawn",
                                hour > 7 & hour < 11 ~ "morning",
                                hour > 10 & hour < 15 ~ "midday", 
                                hour > 14 & hour < 17 ~ "afternoon",
                                hour > 16 & hour < 20 ~ "dusk",
                                TRUE ~ "night")) %>% 
  group_by(date, timeofday) %>% 
  summarize(mean_timeofday_bots = mean(fluxL_umolm2sec, na.rm = TRUE),
            stand = stand) %>% 
  unite("date_timeofday", date:timeofday, sep = ", ", remove = FALSE)

# combine:
means_timeofday <- merge(HFauto_timeofday, HFbots_timeofday, by = "date_timeofday")
```

Plot: by healthy, unhealthy plots

```{r}
legend_title <- "time of day"
means_timeofday$timeofday.x <- factor(means_timeofday$timeofday.x,
                                      levels = c("dawn", "morning", "midday", "afternoon", "dusk", "night"))

means_timeofday %>% 
  ggplot(aes(x=mean_timeofday_auto, y=mean_timeofday_bots, color = timeofday.x, fill = timeofday.x))+
  geom_point(color = "black", pch = 21)+
  geom_smooth(aes(color=timeofday.x), method = "lm", se = FALSE, linewidth = 0.5)+
  scale_fill_manual(legend_title, values = c("white", "lemonchiffon2", "yellow2", "goldenrod3", "steelblue4", "grey26"))+
  scale_color_manual(legend_title, values = c("white", "lemonchiffon2", "yellow2", "goldenrod3", "steelblue4", "grey26"))+
  stat_poly_eq(aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
               formula = y~x,
               parse = TRUE,
               label.x = "right",
               label.y = "bottom",
               )+
  labs(x = "mean flux, HF autochambers", y = "mean flux, Fluxbots")+
  # facet_wrap(~stand.x)+
  theme_dark()

```

Plot: by time of day

```{r}
legend_title <- "plot condition"
means_timeofday$stand.x <- factor(means_timeofday$stand.x,
                                      levels = c("healthy", "unhealthy"))

means_timeofday %>% 
  ggplot(aes(x=mean_timeofday_auto, y=mean_timeofday_bots, color = stand.x, fill = stand.x))+
  geom_point(color = "black", pch = 21)+
  geom_smooth(aes(color=stand.x), method = "lm", se = FALSE, linewidth = 0.5)+
  scale_fill_manual(legend_title, values = c("darkolivegreen3", "lightsalmon1"))+
  scale_color_manual(legend_title, values = c("darkolivegreen3", "lightsalmon1"))+
  stat_poly_eq(aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
               formula = y~x,
               parse = TRUE,
               label.x = "right",
               label.y = "bottom",
               vstep = .1,
               size = 3
               )+
  labs(x = "mean flux, HF autochambers", y = "mean flux, Fluxbots")+
  facet_wrap(~timeofday.x)+
  theme_dark()

```

Plot means of data across time:

```{r}
a <- means_timeofday %>% 
  ggplot(aes(x=timeofday.x, y=mean_timeofday_auto, fill=stand.x))+
  geom_boxplot()+
  geom_jitter()+
  theme_classic2()+
  theme(legend.position = c(.2, .1),
        legend.text = element_text(size = 8), legend.title = element_blank())+
  ylim(0,6)+
  labs(x="", y="average flux by time of day", title = "HF autochambers")+
  scale_fill_manual(legend_title, values = c("darkolivegreen3", "lightsalmon1"))

b <- means_timeofday %>% 
  ggplot(aes(x=timeofday.x, y=mean_timeofday_bots, fill=stand.x))+
  geom_boxplot()+
  theme_classic2()+
  theme(legend.position = c(.2, .1),
        legend.text = element_text(size = 8), legend.title = element_blank())+
  ylim(0,6)+
  labs(x="", y="average flux by time of day", title = "Fluxbots")+
  scale_fill_manual(legend_title, values = c("darkolivegreen4", "lightsalmon3"))

ggarrange(a,b)

```

redo the summary stats with time of day associated:

```{r}
HF_fluxestimates <- HF_fluxestimates %>% 
  mutate(date = as_date(hour_of_obs)) %>% 
  # pull "hour" out:
  mutate(hour = hour(ymd_hms(hour_of_obs))) %>% 
  # assign time of day to each observation:
  mutate(timeofday = case_when(hour > 4 & hour < 8 ~ "dawn",
                                hour > 7 & hour < 11 ~ "morning",
                                hour > 10 & hour < 15 ~ "midday", 
                                hour > 14 & hour < 17 ~ "afternoon",
                                hour > 16 & hour < 20 ~ "dusk",
                                TRUE ~ "night"))

HFauto_fluxestimates <- HFauto_fluxestimates %>% 
  # mutate(date = as_date(hour_of_obs)) %>% 
  # pull "hour" out:
  # mutate(hour = hour(hour_of_obs)) %>% 
  # assign time of day to each observation:
  mutate(timeofday = case_when(hour > 4 & hour < 8 ~ "dawn",
                                hour > 7 & hour < 11 ~ "morning",
                                hour > 10 & hour < 15 ~ "midday", 
                                hour > 14 & hour < 17 ~ "afternoon",
                                hour > 16 & hour < 20 ~ "dusk",
                                TRUE ~ "night"))

# summary stats by time of day:
HF_fluxestimates$timeofday <- factor(HF_fluxestimates$timeofday,
                                      levels = c("dawn", "morning", "midday", "afternoon", "dusk", "night"))
HF_fluxestimates %>%
  filter(fluxL_umolm2sec > -1) %>%
  group_by(timeofday) %>% 
  summarise(mean_flux = mean(fluxL_umolm2sec, na.rm = TRUE),
            min_flux = min(fluxL_umolm2sec, na.rm = TRUE),
            max_flux = max(fluxL_umolm2sec, na.rm = TRUE),
            sd_flux = sd(fluxL_umolm2sec, na.rm = TRUE),
            tenth_quantile = quantile(fluxL_umolm2sec, probs = .1, na.rm = TRUE),
            ninetieth_quantile = quantile(fluxL_umolm2sec, probs = .9, na.rm = TRUE),
            variance = var(fluxL_umolm2sec))

HFauto_fluxestimates$timeofday <- factor(HFauto_fluxestimates$timeofday,
                                      levels = c("dawn", "morning", "midday", "afternoon", "dusk", "night"))
HFauto_fluxestimates %>%
  filter(rs < 15) %>%
  group_by(timeofday) %>% 
  summarise(mean_flux = mean(rs, na.rm = TRUE),
            min_flux = min(rs, na.rm = TRUE),
            max_flux = max(rs, na.rm = TRUE),
            sd_flux = sd(rs, na.rm = TRUE),
            tenth_quantile = quantile(rs, probs = .1, na.rm = TRUE),
            ninetieth_quantile = quantile(rs, probs = .9, na.rm = TRUE),
            variance = var(rs, na.rm = TRUE))
```

Plot non-meaned data across time:

```{r}
c <- HFauto_fluxestimates %>% 
  filter(rs < 15) %>% 
  ggplot(aes(x=timeofday, y=rs, fill=stand))+
  geom_boxplot()+
  theme_classic2()+
  theme(legend.position = c(.2, .9),
        legend.text = element_text(size = 8), legend.title = element_blank())+
  ylim(0,12)+
  labs(x="", y="spread of flux estimates by time of day", title = "HF autochambers")+
  scale_fill_manual(legend_title, values = c("darkolivegreen3", "lightsalmon1"))

d <- HF_fluxestimates %>% 
  filter(fluxL_umolm2sec > 0) %>% 
  ggplot(aes(x=timeofday, y=fluxL_umolm2sec, fill=stand))+
  geom_boxplot()+
  theme_classic2()+
  theme(legend.position = c(.2, .9),
        legend.text = element_text(size = 8), legend.title = element_blank())+
  ylim(0,12)+
  labs(x="", y="spread of flux estimates by time of day", title = "Fluxbots")+
  scale_fill_manual(legend_title, values = c("darkolivegreen4", "lightsalmon3"))

ggarrange(c,d)

```

Simple statistics comparing healthy, unhealthy plots:

```{r}
# using library rstatix for tidy-compatible testing:

# HF AUTOCHAMBERS
t <- HFauto_fluxestimates %>% 
  filter(rs < 15) %>% 
  t_test(rs ~ stand, ref.group = "healthy", detailed = TRUE)
t
# healthy sample size: 6121, healthy mean: 3.048552
# unhealthy sample size: 7750, unhealthy mean: 2.33636
# statistic = 34.98524	
# df = 8264.074		
# p = 3.45e-250

# FLUXBOTS
t2 <- HF_fluxestimates %>% 
  filter(fluxL_umolm2sec > 0) %>% 
  t_test(fluxL_umolm2sec ~ stand, ref.group = "healthy", detailed = TRUE)
t2
# healthy sample size: 3639, healthy mean: 2.33272
# unhealthy sample size: 3366, unhealthy mean: 2.541244
# statistic = -6.67173		
# df = 6674.28	
# p = 2.73e-11	
```

```{r shit}


HF_fluxestimates <- HF_fluxestimates %>%
  filter(fluxQ_umolm2sec >= 0)  %>%   filter(fluxL_umolm2sec >= 0) 
source("filter_iqr.R")
HF_fluxestimates<- HF_fluxestimates %>% filter_iqr('fluxQ_umolm2sec') %>% filter_iqr('fluxL_umolm2sec') 


HFauto_fluxestimates <- HFauto_fluxestimates %>%
  filter(rs >= 0) 
source("filter_iqr.R")
HFauto_fluxestimates<- HFauto_fluxestimates %>% filter_iqr('rs') 

ggplot(data=HF_fluxestimates, aes(x=ymd_hms(hour_of_obs), y=final_flux_umolm2sec))+
  geom_point(alpha=0.2)+
  geom_smooth(method = "loess", span = 0.01, color = "blue", se=F)+
  theme_classic()+
  facet_wrap(~stand, ncol=1)

ggplot(data = HFauto_fluxestimates, aes(x = ymd_hms(date_time), y = rs)) +
  geom_point(alpha=0.2)+
  geom_smooth(method = "loess", span = 0.01, color = "blue", se=F)+
  theme_classic()+
  facet_wrap(~stand, ncol=1)
  









# Summarize daily means by stand
fluxbot_daily_means <- HF_fluxestimates %>%
  group_by(stand, date) %>%
  summarize(daily_mean_flux = mean(fluxL_umolm2sec, na.rm = TRUE))

# Summarize daily means by stand
autochamber_daily_means <- HFauto_fluxestimates %>%
  group_by(stand, date) %>%
  summarize(daily_mean_rs = mean(rs, na.rm = TRUE))


merged_data <- full_join(fluxbot_daily_means, autochamber_daily_means, by = c("stand", "date"))

ggplot(merged_data, aes(x=daily_mean_rs, y=daily_mean_flux))+
  geom_point(aes(color=stand))+
  geom_smooth(method="lm")+
  stat_poly_eq(aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
               formula = y~x,
               parse = TRUE,
               label.x = "right",
               label.y = "top")


summary(lm(daily_mean_flux~daily_mean_rs*stand, data=merged_data))














# Summarize daily means by stand
fluxbot_hourly_means <- HF_fluxestimates %>%
  group_by(stand, hour_of_obs) %>%
  summarize(hourly_mean_flux = mean(fluxL_umolm2sec, na.rm = TRUE))


HFauto_fluxestimates$hour_of_obs<-lubridate::floor_date(lubridate::ymd_hms(HFauto_fluxestimates$datetime), "hour")

# Summarize daily means by stand
autochamber_hourly_means <- HFauto_fluxestimates %>%
  group_by(stand, hour_of_obs) %>%
  summarize(hourly_mean_rs = mean(rs, na.rm = TRUE))


#merged_data <- full_join(fluxbot_hourly_means, autochamber_hourly_means, by = c("stand", "hour_of_obs"))
# 
# ggplot(merged_data, aes(x=hourly_mean_rs, y=hourly_mean_flux))+
#   geom_point(aes(color=stand))+
#   geom_smooth(method="lm")+
#   #facet_wrap(~stand)+
#   geom_abline(intercept = 0, slope=1)+
#    stat_poly_eq(aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
#                formula = y~x,
#                parse = TRUE,
#                label.x = "right",
#                label.y = "top")
#   


#summary(lm(daily_mean_flux~daily_mean_rs*stand, data=merged_data))




ggplot(data=HFauto_fluxestimates,
       aes(x=rs, color=stand))+
  geom_histogram()+
  facet_wrap(~chamber)


ggplot(data=HFauto_fluxestimates,
       aes(x=rs, y=as.factor(chamber), color=stand))+
  geom_jitter(alpha=0.3)+
  geom_boxplot(outliers = FALSE, color="black", fill=NA)

library(ggridges)
ggplot(data=HFauto_fluxestimates,
       aes(x=rs, y=as.factor(chamber), color=stand))+
  geom_density_ridges()


summary(aov(rs~stand*as.factor(chamber), data=HFauto_fluxestimates))
summary(aov(fluxL_umolm2sec~stand*as.factor(id), data=HF_fluxestimates))




ggplot(data=HFauto_fluxestimates,
       aes(x=rs, fill=stand))+
  geom_density(alpha=0.5)

ggplot(data=HF_fluxestimates,
       aes(x=ymd_hms(hour_of_obs), y=fluxL_umolm2sec))+
  geom_point()+
  facet_wrap(~id)

ggplot(data=HFauto_fluxestimates,
       aes(x=ymd_hms(hour_of_obs), y=rs))+
  geom_point()+
  facet_wrap(~chamber)





HF_fluxestimates <- HF_fluxestimates %>%
  arrange(stand)

ggplot(data=HF_fluxestimates,
       aes(x=fluxL_umolm2sec, color=stand))+
  geom_histogram()+
  facet_wrap(~id)

ggplot(data=HF_fluxestimates,
       aes(x=fluxL_umolm2sec, y=id, color=stand))+
  geom_jitter(alpha=0.3)+
  geom_boxplot(outliers = FALSE, color="black", fill=NA)


ggplot(data=HF_fluxestimates,
       aes(x=fluxL_umolm2sec, y=id, color=stand))+
  geom_density_ridges()


# merged_data_long <- merged_data %>%
#   pivot_longer(cols = c(hourly_mean_flux, hourly_mean_rs), 
#                names_to = "measurement_type", 
#                values_to = "value") %>%
#   mutate(measurement_type = recode(measurement_type,
#                                    'hourly_mean_flux' = 'fluxbot',
#                                    'hourly_mean_rs' = 'autochamber'))
# 
# ggplot(merged_data)+
#   geom_density(aes(x=hourly_mean_flux), fill="#00356B", alpha=0.5)+
#   geom_density(aes(x=hourly_mean_rs), fill="forestgreen", alpha=0.5)+
#   facet_wrap(~stand)
```
